---
title:  "DIA proteomics (from DIA-NN) analysis with an MS-EmpiRe"
author: "BS"
date:   "09/10/2022"
output:  github_document
---

initially directory should only contain the main output of DIA-NN and contaminats fasta file from MaxQuant (latter only necessary if it was used during DIA-NN search)

pepquantify - https://github.com/bshashikadze/pepquantify
Ms-EmpiRe   - https://github.com/zimmerlab/MS-EmpiRe

### load libraries
```{r message=FALSE, warning=FALSE}
library      (msEmpiRe)    # differential abundance analysis
library      (seqinr)      # filtering for contaminants
library      (tidyverse)   # data manipulation
library      (pepquantify) # prepare data for MS-EmpiRe analysis
```


### read the main output of DIA-NN
implicit reading is necessary only if any modifications other than supported by pepquantify package is necessary, e.g. filtering for contaminats
```{r message=FALSE, warning=FALSE}
raw_diann <- read.delim("DIA-NN_output_precursors.tsv", sep = "\t", header = T) #can be downloaded from github
```


### remove contaminants (contaminants fasta file from MaxQuant common contaminants)
```{r}
contamintants           <- read.fasta("contaminants.fasta")
contaminant.names       <- getName(contamintants) 
raw_diann_filtered      <- raw_diann %>% 
  filter(!str_detect(Protein.Group, str_c(contaminant.names, collapse="|")))

# move original file to the separate folder (this is because "pepquantify" will read automatically largest tsv file so it is necessary to leave only filtered data in the main directory)
dir.create("original")
file.copy(from = paste0(getwd(), "/DIA-NN_output_precursors.tsv"),
          to   = paste0(getwd(), "/original/DIA-NN_output_precursors.tsv"))
unlink("DIA-NN_output_precursors.tsv")

# save contaminants-removed data
write.table(raw_diann_filtered, "MIDY_Lung_DIA.tsv", quote = F, sep = "\t", row.names = F)
```



```{r}
msempire_calculation <- function(data, data2 = data_raw, seed=1234, fc_threshold = 1.5) {
  
  require(magrittr)
  # read the data in the expressionset format and perform msempire normalization and quantification  
  # (https://github.com/zimmerlab/MS-EmpiRe/blob/master/example.R)
  msempiredata  <- msEmpiRe::read.standard(data[[1]], data[[2]],
                                            prot.id.generator = function(pep) unlist(strsplit(pep, "\\.[0-9]*$"))[1],
                                            signal_pattern="Intensity.*")
  
  # msempire calculations
  set.seed(seed = seed)
  msempire_results <- msempiredata  %>%
    msEmpiRe::normalize() %>%
    msEmpiRe::de.ana() %>%
    write.table(paste0(data[[3]], "/msempire_results_raw.txt"), sep = "\t", row.names = F)
  
  
# tidy results (pepquantify package)
pepquantify::resultstidy(data, data2,  fc_threshold = fc_threshold)}
```


```{r message=FALSE, warning=FALSE}
data_raw <- pepquantify::read_diann(experimental_library = T)
```


```{r message=FALSE, warning=FALSE}
msempire_data <- pepquantify::pepquantify_funs(data_raw, condition1 = "MIDY", condition2 = "WT", imputation = TRUE)
msempire_calculation(msempire_data, fc_threshold = 1.5)
```






