---
title: "Multi (uni-) variate analysis of lipidomics data"
author: "B.Shashikadze"
date: "21/08/2022"
output: github_document
---



### load libraries
```{r message=FALSE}
library(tidyverse) 
library(ggrepel)  
library(ropls)    
library(ggpubr)
```


### load omics data and clean-up
Normal wide-table format with compound names as rows followed by the samples as columns
```{r}
data_raw <- read.delim("lipidomics_raw.txt", sep = "\t", header = T, check.names = F) %>% 
  filter(if_all(-contains("Compound"), ~ (.x >= 0))) %>% 
  rename_all(~str_replace(., "MIDY", "")) %>% 
  rename_all(~str_replace(., "WT", ""))

# class of each column
sapply(data_raw, class)

# convert all but first column to numeric
data_raw[ , 2:10] <- apply(data_raw[ , 2:10], 2,            
                    function(x) as.numeric(as.character(x)))
```

### define groups
```{r}
Bioreplicate <- names(data_raw)[-1]
Condition <- Bioreplicate
Groups <- data.frame(Bioreplicate, Condition)
write.table(Groups, "conditions.txt", row.names = F, quote = F, sep = "\t")
cat("conditions file was generated rename the file as Groups_modified, and modify the second column according to groups")
```


### read modified Groups file
```{r}
Groups <- read.delim("conditions_modified.txt", header = T, sep = "\t")
Groups$Bioreplicate  <- as.character(Groups$Bioreplicate)
```



### function to calculate p-values and fold changes
```{r}
statistic_function <- function(data, conditions_data, condition, parametric = TRUE, first_group, padjmethod = "BH", id_name, values_log) {

  if (values_log == TRUE) {
    data_long <- data %>% 
    pivot_longer(names_to = "Bioreplicate", values_to = "Intensity", -!!as.symbol(id_name)) %>%
    left_join(conditions_data)
  }
  
  if (values_log == FALSE) {
    data_long <- data %>% 
    pivot_longer(names_to = "Bioreplicate", values_to = "Intensity", -!!as.symbol(id_name)) %>%
    mutate(Intensity = log2(Intensity)) %>% 
    left_join(conditions_data)
  }
  
  #fold-change calculation    
    data_fc <- data_long %>% 
    group_by(!!as.symbol(id_name), !!as.symbol(condition)) %>% 
    summarise(grp_mean = mean(Intensity)) %>% 
    ungroup() 
  
    
  if (first_group %in% unique(data_fc[[2]][2])) {
    data_fc <- data_fc %>% 
     select(-all_of(condition)) %>% 
     group_by(!!as.symbol(id_name)) %>% 
     mutate(l2fc=grp_mean-lag(grp_mean)) %>% 
     select(-grp_mean)  %>% 
     drop_na()
     }
  
   else{
    data_fc <- data_fc  %>% 
    select(-all_of(condition)) %>%  
    group_by(!!as.symbol(id_name))  %>% 
    mutate(l2fc=grp_mean-lead(grp_mean)) %>% 
    select(-grp_mean)   %>% 
    drop_na()
     }
    
    cat(paste("positive fold change means up in", first_group, sep=" "))
  
  # statistics calculation 
  if (parametric == TRUE) {
   
  data_p <- data_long  %>% 
  group_by(!!as.symbol(id_name))  %>% 
  summarise(t.test_p.val = t.test(Intensity ~ !!as.symbol(condition), var.equal=F)$p.value) %>% 
  left_join(data_fc)  
  data_p$p.adj <- p.adjust(data_p$t.test_p.val, method = padjmethod)
    
  cat(paste("p-values were adjusted using the", padjmethod, "method", sep=" "))
    
  return(data_p)}
  
  else {
  
  data_p <- data_long %>% 
  group_by(!!as.symbol(id_name))  %>% 
  summarise(wilk.test_p.val = wilcox.test(Intensity ~ !!as.symbol(condition), exact = F)$p.value) %>% 
  left_join(data_fc)
  
  data_p$p.adj <- p.adjust(data_p$wilk.test_p.val, method = padjmethod)
  
  cat(paste("p-values were adjusted using the", padjmethod, "method", sep=" "))
    
    return(data_p)
    }}

statistics_data <- statistic_function(data_raw, Groups, condition = "Condition", parametric = TRUE, first_group = "MIDY", padjmethod = "none", id_name = "Compound", values_log = F)
```

## multivariate analysis

### prepare data for the multivariate analysis
```{r}
multivariate_prep_function <- function(data){
  data <- data %>% 
          column_to_rownames(var="Compound") %>% 
          log2() %>% 
          t() %>% 
          as.data.frame()
}
data_multi <- multivariate_prep_function(data_raw)
```


### principal component analysis (calculations)
```{r}
pca_function <- function(data) {
  PCA <- prcomp(data)
  pca.data <- data.frame(Bioreplicate=rownames(PCA$x),
  X=PCA$x[,1],
  Y=PCA$x[,2])
  pca.var <- PCA$sdev^2
  pca.var.per <- round(pca.var/sum(pca.var)*100,1)
  pca.data <- pca.data %>% 
    left_join(Groups)
  data_pca <- list()
  data_pca[[1]] <- pca.data
  data_pca[[2]] <- pca.var.per
  return(data_pca)
}
data_pca <- pca_function(data_multi)
```


### principal component analysis (plotting)
```{r}
pcaplot <- ggplot(data=data_pca[[1]], aes(x=X, y=Y, label=Bioreplicate, fill= Condition))+
geom_point(size = 3.5, shape = 21)+
scale_fill_manual(values=c('MIDY' = "firebrick3", 'WT' = "#0072B2"))+
geom_text_repel(colour="black", size = 3.5, box.padding = 0.3)+
xlab(paste("PC axis 1 - ", data_pca[[2]][1], "%", sep=""))+
ylab(paste("PC axis 2 - ", data_pca[[2]][2], "%", sep=""))+
geom_hline(yintercept = 0, linetype = "dashed")+
geom_vline(xintercept = 0, linetype = "dashed")+
theme_bw() + theme(panel.border = element_rect(size = 1),
                   axis.title = element_text(size = 10, colour="black"), 
                   axis.text.x = element_text(size=10, colour="black"),
                   axis.text.y = element_text(size = 10, colour="black"),
panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
                      theme(legend.position = "top", legend.box.spacing = unit(0.5, 'mm'), legend.title = element_blank(), legend.text = element_text(size = 10))
```


### Orthogonal Projections to Latent Structures Discriminant Analysis (OPLS-DA) (calculations)
```{r}
oplsda_function <- function (data, scaling = "pareto", n_perm = 200, 
                             n_crossval = 9, 
                             vip_thresh = 1.5)  {
  Grouping     <- Groups[, "Condition"]
  oplsda_model <- opls(data, Grouping, predI = 1, 
                      orthoI = NA, permI = n_perm, 
                      scaleC = scaling, 
                      crossvalI = n_crossval, subset = NULL) 
          p1          <- round(oplsda_model@modelDF$`R2X(cum)`[1]*100)
    loadings          <- getLoadingMN(oplsda_model)
  oplsda_vip          <- oplsda_model@vipVn %>% 
    as.data.frame() %>% 
    rownames_to_column() %>% 
    rename(Compound = rowname, VIP = 2) %>% 
    left_join(statistics_data) %>% 
    mutate(VIP_sig = case_when(VIP > vip_thresh ~ "+")) %>% 
    mutate(Regulation = case_when(
      l2fc > 0 & VIP_sig == "+" ~ "upregulated",
      l2fc < 0 & VIP_sig == "+" ~ "downregulated",
      TRUE ~ "n.s."
    ))  
  oplsda_vip$Compound <- factor(oplsda_vip$Compound, levels = oplsda_vip$Compound[order(oplsda_vip$VIP)])
  oplsda_pred         <- as.data.frame(oplsda_model@scoreMN)
  oplsda_ortho        <- as.data.frame(oplsda_model@orthoScoreMN)
  oplsda_components   <- cbind(oplsda_pred, oplsda_ortho, Groups)
  data_oplsda         <- list()
  data_oplsda[[1]]    <- oplsda_components
  data_oplsda[[2]]    <- oplsda_vip
  data_oplsda[[3]]    <- p1
  return(data_oplsda)
}
set.seed(123456) 
data_oplsda <- oplsda_function(data_multi, scaling = "pareto", n_perm = 200, n_crossval = 9, vip_thresh = 1.5)
```



### variance importance in projection (plotting)
```{r}
vipplot <- ggplot(data_oplsda[[2]] %>% 
                  filter(VIP_sig == "+"), aes(x = VIP, y= Compound, fill = Regulation)) +
geom_point(size = 3.5, shape = 21) +
theme_bw()+
scale_fill_manual(values = c("firebrick3", "#0072B2"))+
theme(panel.border = element_rect(size = 1), panel.grid.major = element_line(), panel.grid.minor = element_blank(), panel.background = element_blank(),
                            axis.text.x = element_text(colour = "black", size = 10),
                            axis.title.x = element_text(size= 10),
                            axis.title.y = element_blank(),
                            axis.text.y = element_text(size = 10, colour = "black"))+
xlab("VIP score")+
ggtitle("VIP plot") +
theme(plot.title = element_blank()) +
theme(legend.position = "top", legend.box.spacing = unit(0.5, 'mm'), 
      legend.title = element_blank(), 
      legend.text = element_text(size = 10))
```



### volcano plot highlighting compounds with a large VIP scores
```{r}
volcanoplot <- ggplot(data_oplsda[[2]], mapping = aes(x = l2fc, y = t.test_p.val, fill=Regulation, label = Compound, alpha = Regulation))+
geom_point(aes(shape =Regulation, size = Regulation))+
scale_shape_manual(values = c (n.s. = 16, downregulated =21, upregulated =21))+
scale_y_reverse(breaks = c(1,0.85,0.70,0.55,0.40,0.25,0.15, 0.05), limits = c(1, 0))+
scale_size_manual(values=c (n.s. = 2, downregulated =2, upregulated =2))+
scale_fill_manual(values=c(n.s. = "#4a4949", downregulated = "firebrick3", upregulated = "#0072B2"))+
scale_alpha_manual(values= c(n.s. = 0.7, "Downregulated"= 1, "Upregulated"= 1))+
geom_text_repel(
  data = subset(data_oplsda[[2]], Regulation != "n.s."),
  aes(label = Compound),
  size = 2.2,
  color = "black",
  box.padding = unit(0.40, "lines"),
  point.padding = unit(0.40, "lines"),
  max.overlaps = Inf,
  alpha = 1
  )+
theme_bw()+
theme(panel.border = element_rect(size = 1), panel.grid.major = element_line(), panel.grid.minor = element_blank(), panel.background = element_blank(),
                            axis.text.x = element_text(colour = "black", size = 10),
                            axis.title.x = element_text(size= 10),
                            axis.title.y =element_text(colour = "black", size = 10),
                            axis.text.y = element_text(size = 10, colour = "black"))+
xlab("log2 fold change (MIDY/WT)")+
ylab("P-value")+
                      theme(plot.title = element_blank()) +
                      theme(legend.position = "top", 
                            legend.box.spacing = unit(0.5, 'mm'), 
                            legend.title = element_blank(), 
                            legend.text = element_text(size = 10))+
scale_x_continuous(breaks = c(-1,0,1), limits = c(-1.8, 1.8))
```



### Orthogonal Projections to Latent Structures Discriminant Analysis (OPLS-DA) (plotting)
```{r}
oplsdaplot <- ggplot(data=data_oplsda[[1]], aes(x=p1, y=o1, label=Bioreplicate, fill = Condition))+
geom_point(size = 3.5, shape =21)+
scale_fill_manual(values=c(MIDY = "firebrick3", WT = "#0072B2"))+
geom_text_repel(colour="black", size = 3.5, box.padding = 0.3)+
xlab(paste("OPLS-DA axis 1 - ", data_oplsda[[3]], "%", sep=""))+
ylab("OPLS-DA axis 2")+
geom_hline(yintercept = 0, linetype = "dashed")+
geom_vline(xintercept = 0, linetype = "dashed")+
theme_bw() + 
  theme(panel.border = element_rect(size = 1),
                   axis.title = element_text(size = 10, colour="black"), 
                   axis.text.x = element_text(size=10, colour="black"), 
                   axis.text.y = element_text(size = 10, colour="black"),
panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
theme(legend.position = "top", legend.box.spacing = unit(0.5, 'mm'), 
      legend.title = element_blank(), 
      legend.text = element_text(size = 10))
```


# combine plots
```{r}
# row 1
P1 <- ggarrange(pcaplot, oplsdaplot, widths = c(3.4, 3.4), heights = c(3.4, 3.4),
          labels = c("a", "b"), font.label = list(size = 22, face = 'bold'), 
          ncol = 2, nrow = 1,  legend = "bottom",
          common.legend = T)
# row 2
P2 <- ggarrange(vipplot, volcanoplot, widths = c(3.4, 3.4), heights = c(3.4, 3.4),
          labels = c("c", "d"), font.label = list(size = 22, face = 'bold'),
          ncol = 2, nrow = 1,  legend = "bottom",
          common.legend = T)
# combine all and save
ggarrange(P1,P2, 
          ncol = 1, nrow = 2, widths = c(6.8, 6.8), heights = c(6.8, 6.8))
ggsave("lipidomics.svg", height = 6.8, width = 6.8)
```




# save
```{r}
lipidomics_stat <- data_oplsda[[2]] %>% 
  arrange(desc(VIP)) %>% 
  select(1:4) %>% 
  write.table("lipidomics_stat.txt", sep = "\t", row.names = F, quote = F)

save.image()
```